---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

The assumptions below were built on the brfss2013 dataset, which is a big dataset collected by surveying random people accross the united states. Because of that, this assumptions can be generalized for at least all the United States.

We can not make assumptions of causality, we can only talk about correlation so all the hypothesis below are referring to correlation by the word "relationship".

* * *

## Part 2: Research questions

**Research quesion 1: Does the sleep time follow a normal distrubution ?**

**Research quesion 2: Is there a difference for smokers?**

**Research quesion 3: Is there a relationship between income and poor health?**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}

brfss2013 %>% ggplot(aes(x=sleptim1)) + geom_histogram(binwidth = 1)  + xlim(2,12) + labs(title="Histogram of sleeptime",x="Number of hours slept per day") 

```

It looks like it has a distrubution that looks a lot like a normal distrubution but a little left-skewed.

This can be generalized to the whole united states.



```{r}

mean(brfss2013$sleptim1,na.rm = TRUE)

```

An average person sleeps 7.052 hours a day.

```{r}

sd(brfss2013$sleptim1,na.rm = TRUE)

```

and the standard deviation of sleep time per day is 1.604.

So we can create a normal distribution with this values and compare it with the original.

```{r}

hours_sleep <- seq(2,12,by=1)


percentage <- dnorm(hours_sleep,mean=mean(brfss2013$sleptim1,na.rm=TRUE),sd=sd(brfss2013$sleptim1,na.rm=TRUE))

plot(hours_sleep,percentage)

```


**The percentage value here stands for the percentage of people sleeps given number of hours.** 

And, any assumption based on this distribution would be very close to the reality in the United States.



**Research quesion 2:**

```{r}
smokers <- brfss2013 %>% filter(usenow3 != "Not at all")
```

First, we filter the non-smokers from our data to work especially on the smokers.


```{r}
smokers %>% ggplot(aes(x=sleptim1)) + geom_histogram(binwidth=1) + xlim(2,12) + labs(title="Histogram of sleeptime (smokers)",x="Number of hours slept per day") 
```

By comparing the this plot and the one we created in the beginning, we can see that there is a little difference between the sleep times of smokers in general. Their distribution seems less left skewed. We can say that a big portion of people sleep less than the average are smokers.

To be more clear, lets see the average and the standard deviation, then generate a normal distribution with those variables.

```{r}

mean(smokers$sleptim1,na.rm = TRUE)

```

Average sleep time of a smoker is 6.876 (general average was 7.05)

```{r}

sd(smokers$sleptim1,na.rm = TRUE)

```

Standard deviation is 1.589 (general standard deviation was 1.604)


```{r}
percentage <- dnorm(hours_sleep,mean=mean(smokers$sleptim1,na.rm=TRUE),sd=sd(smokers$sleptim1,na.rm=TRUE))

plot(hours_sleep,percentage)

```

As it can be easier to see from this plot, the difference between the smokers and the general is not too great.


**Research question 3:**

```{r}
brfss2013 %>% filter(!is.na(income2)) %>% ggplot(aes(x=income2,y=poorhlth)) + geom_boxplot() + ylim(0,30) + labs(title="Boxplot of income(x axis) and poor health(y axis)" , x="Income Level" , y="number of days felt poor health in the last month")
```

From the plot above, we can easily see that people with less income tend to have poor health more than the people with higher income. Especially below $15 000 threshold. After that, the median value gets closer to the zero and the 75th percentile falls down.

